---
title: "Practicum Fundamentals"
author: "Abdul Moid"
date: "2023-06-24"
output: html_document
---

```{r code, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)

# Load required libraries
library(readxl)
library(dplyr)
library(ggplot2)
library(openxlsx)

# Read the Excel file using read.xlsx()
survey_data <- read.xlsx("C:/Users/mfhus/Downloads/survey.xlsx", sheet = 1, detectDates = TRUE)

# Convert columns to appropriate types
survey_data$Timestamp <- as.POSIXct(survey_data$Timestamp)  # Convert to datetime
survey_data$Age <- as.integer(survey_data$Age)  # Convert to integer
survey_data$Gender <- as.factor(survey_data$Gender)  # Convert to factor

# Examine the structure of the dataset
str(survey_data)

# Summary statistics of the dataset
summary(survey_data)

# 1.1 Missing Values

# Check for missing values
colSums(is.na(survey_data))

# Remove rows with missing values
survey_data <- survey_data %>% na.omit()


# 1.2 Incorrect Data Types

# Check data types of variables
str(survey_data)

# Convert variables to appropriate data types
survey_data$Timestamp <- as.POSIXct(survey_data$Timestamp)  # Convert to datetime
survey_data$Age <- as.integer(survey_data$Age)  # Convert to integer
survey_data$Gender <- as.factor(survey_data$Gender)  # Convert to factor
survey_data$Country <- as.factor(survey_data$Country)  # Convert to factor

# Check data types after conversion
str(survey_data)


# 1.3 Spelling Errors

# Identify unique values and correct any spelling errors
unique(survey_data$Gender)

# Correct spelling errors
survey_data$Gender <- recode(survey_data$Gender, "F" = "Female", "M" = "Male")

# Check unique values again after correction
unique(survey_data$Gender)


```{r code2, echo=TRUE}

# 2. Combining Data Files

# Read the second sheet from the Excel file
survey_data_2 <- read_excel("C:/Users/mfhus/Downloads/survey.xlsx", sheet = 2)

# Combine the two datasets
combined_data <- bind_rows(survey_data, survey_data_2)


```{r code3, echo=TRUE}

# 3. Visualizations

# Histogram of Age variable
ggplot(combined_data, aes(x = Age)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
  labs(x = "Age", y = "Count", title = "Distribution of Age")

# Boxplot of Country variable
ggplot(combined_data, aes(x = "Age", y = Country)) +
  geom_boxplot(fill = "lightgreen", color = "black") +
  labs(x = "Age", y = "Country", title = "Boxplot of Country")






